---
title: 'Lab 03 - Regular Expressions'
output:
  html_document:
    theme: readable
    df_print: paged
    highlight: tango
    toc: true
    self_contained: false
    number_sections: false
    css: labs.css
    include:
      after_body: footer.html    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set( echo = TRUE, message=F, warning=F, fig.width=10 )
library( dplyr )
library( pander )
```



# Data

This lab utilizes a [free Kaggle dataset](https://www.kaggle.com/dorianlazar/medium-articles-dataset/notebooks) describing the popularity of articles on the blogging platform Medium. We will use the following fields: 

* **title** - the blog title 
* **subtitle** - blog subtitle (optional)
* **claps** - the number of "likes" each article received 
* **reading_time** - time (in minutes) it takes to reach each article 
* **publication** - Medium community in which the blog was published 
* **date** - date of publication in style YYYY-MM-DD 


```{r}
# URL <- "https://www.dropbox.com/s/tizgdsat2mziod6/medium-data-utf8.csv?dl=1"
URL <- "https://raw.githubusercontent.com/DS4PS/cpp-527-fall-2020/master/labs/data/medium-data-utf8.csv"
d <- read.csv( URL )

preview.these <- c("title", "subtitle", "claps", "reading_time", "publication", "date")
head( d[preview.these] ) %>% pander()
```




## Text Pre-Processing Steps

The article title text is fairly clean and needs little pre-processing, but there are a couple of important issues that need to be addressed. 

1. Remove HTML tags 
2. Remove "hair space" dash marks

Some titles include HTML tags for emphasis. They will always occur at the start and end of the titles: 

```{r, eval=F}
"<strong class=\"markup--strong markup--h3-strong\">Why 70% of Entrepreneurs Are Ending Their Pitches the Wrong Way</strong>"
"<strong class=\"markup--strong markup--h3-strong\">Tackle Negative Core Beliefs</strong>"
```

There are some stylized dashes in the text that are book-ended with "hair space" characters, which are thinner-than-usual spaces. 

In the UTF-8 text encoding they come out as:

```
<U+200A>—<U+200A>
```

Since they are not actual words, remove all of these hairspace-dashes from the text. 



## Create a Performance Measure 

We can track article performance in many ways. This dataset contains "claps", which are the equivalent of a thumbs-up or like on other platforms. 

It also contains a responses category, but that is a little more complicated to interpret because comment might be asking for clarification because the article was poorly-written or arguing a counter-point becaue the article was controversial. 

We will use **claps** to measure article popularity, but note the skew in the raw measure:

```{r, fig.width=10}
hist( d$clap, main="Raw Clap Count",
      xlim=c(1,950), breaks=5000, col="gray20", border="white" )
hist( log10(d$clap+1), main="Logged Clap Score",
      col="gray20", border="white", breaks=100 )
```

Create a new **clap score** to measure performance using the following log transformation:

```{r}
d$clap.score <- log10( d$clap + 1 )
```




# Lab Questions 





## Q1 - Comparing Title Styles

Which title style consistently gets the most claps (clap score)? 

**Power Lists:**

* "Six Recommendations for Aspiring Data Scientists"  
* "13 Essential Newsletters for Data Scientists: Remastered"  
* "7 Machine Learning lessons that stuck with me this year"  

**How to guides:** 

* "How to measure progress with outcomes" 
* "How We Dumped A/B Testing For Machine Learning" 
* "How To Be Productive Without Being Hard on Yourself" 

**Something Colon Something:**

* "Business Checkup: 10 Ways to Improve your 2020 Communications" 
* "ReFocus: Making Out-of-Focus Microscopy Images In-Focus Again" 
* "Review: SqueezeNet (Image Classification)"

**Questions:**

* "Why Do Financial Models Break?"
* "What Is Econometrics?"
* "Can AI Robots Rebel Against Us, as Shown in Films?"
* "Can You Invest Like a Top Fund Manager?" 
* "So you’re building a “Superhuman of X”?"

**Other:**

All other styles not included in these. 

Use regular expressions in order to identify all of the titles that belong in each category. 

```{r, eval=F}

# grepl returns TRUE for all cases that 
# match the expression and FALSE otherwise 

category.members <- grepl( "expression", d$cleaned.titles )
```


**Q1-A:**What is the average performance (clap score) of each type of title? 

**Q1-B:**Which style of title performs the best overall? 



## Q2 - Content Analysis 


Split each title into a group of distinct words. 

**Q2-A:** What are the 25 most common words used in the titles? 

**Q2-B:** What is the most common word used at the beginning of a title? 

**Q2-C:** What is the most common word used at the end of a title? 

*Hint: be sure to convert all words to the same case so that variants like "How" and "how" are treated the same.* 

```{r,eval=F}
x <- tolower( x )
```



**CHALLENGE QUESTION:** 

Find a package in R that will tabulate sentiment scores for a set of text. 

These packages will return counts or percentages of words with positive salience (e.g. happiness, joy, laughed) and counts or percentages with negative salience (e.g. hate, disdain, smirk). 

Which titles perform better? Those with high positive salience? Or those with high negative salience? 





## Q3 - Length of Title 

Does the length of the title matter? 



**Q3-A:** Do titles with more words get better clap scores? 

Count the number of unique words in each title. 

Then examine the clap score as a function of word count by regressing the clap score onto word count. 

Visualize the relationship using a box and whisker plot: 

```{r}
x <- sample( 1:10, size=1000, replace=TRUE )
y <- 3*x - 0.5*x*x + rnorm(1000,0,5)
f <- factor(x)

# plot( factor, numeric.vector )
plot( f, y )
```


**Q3-B:** Do longer titles get better clap scores? 

Repeat the analysis from Q3-A but this time measureing title length by the number of characters in the title. 



## Challenge Question 

Put all of the steps above together into a single regression model:

* Title style 
* Title length 
* Large proportion of 100 most common words 

Add some additional controls: 

* Presence or absence of a subtitle 
* Length of the article (time it takes to read) 
* Months since it was published 
* The Medium community in which the blog appears: 

```{r}
table( d$publication ) %>% pander()
```

You are also free to add additional features of the titles that you create with regular expressions. 


**Q: Which factors best predict article performance overall?**




<br>
<br>
<br>

# Submission Instructions

When you have completed your assignment, knit your RMD file to generate your rendered HTML file. Platforms like BlackBoard and Canvas often disallow you from submitting HTML files when there is embedded computer code, so create a zipped folder with both the RMD and HTML files.

Login to Canvas at <http://canvas.asu.edu> and navigate to the assignments tab in the course repository. Upload your zipped folder to the appropriate lab submission link.

Remember to:

* name your files according to the convention: **Lab-##-LastName.Rmd**
* show your solution, include your code.
* do not print excessive output (like a full data set).
* follow appropriate style guidelines (spaces between arguments, etc.).


<br>
<br>

----

<br>
<br>



**Notes on Knitting**

Note that when you knit a file, it starts from a blank slate. You might have packages loaded or datasets active on your local machine, so you can run code chunks fine. But when you knit you might get errors that functions cannot be located or datasets don't exist. Be sure that you have included chunks to load these in your RMD file.

Your RMD file will not knit if you have errors in your code. If you get stuck on a question, just add `eval=F` to the code chunk and it will be ignored when you knit your file. That way I can give you credit for attempting the question and provide guidance on fixing the problem.


**Markdown Trouble?**

If you are having problems with your RMD file, visit the [**RMD File Styles and Knitting Tips**](https://ds4ps.org/cpp-526-spr-2020/labs/r-markdown-files.html) manual.



<br>
<br>

----

<br>
<br>




<style>
blockquote {
    padding: 11px 22px;
    margin: 0 0 22px;
    font-size: 18px;
    border-left: 5px solid lightgray;
}
</style>
